Repairing cooperation through a cognitive intervention in the repeated Trust Game

### 2023-06-07 article.Rmd
started changes

```{r load-packages05, include = FALSE}
library(papaja)
library(kableExtra)
require(knitr)
require(citr)
require(bookdown)

@@

At the core of social interaction is figuring out the goals, intentions, and decision-making process of the interaction partner. This inference is however fraught with uncertainty, as we cannot reasonably observe these features without prior knowledge of the person. Absent a history of interaction, one needs to decide whether to take the risk of trusting others. The challenge of trust is that it is by construction a risky endeavour. For example, if we deem a person trustworthy, we might decide to take the risk of investing in the relationship hoping for a collaborative outcome. If we misplace our trust, this can come at a high cost to us. Not trusting others is also risky since opportunities for cooperation may be foregone.

@@

Evidence from the literature emphasises the importance of social trust in determining why some people fare better than others physically and mentally [@giordano_trust_2016 ; @meng_multilevel_2014]. It affects the health status of individuals through reinforcing social support networks, maintaining community norms and facilitating collective action. Research into the determinants of psychopathology has linked trust-based constructs to the emergence of mental health disorders. @fonagy_role_2014 identified epistemic trust, or the belief in the authenticity and personal relevance of interpersonally transmitted knowledge, as an important function of early attachment relationships. It allows the individual receiving social information to let go of their natural self-protective vigilance, which can become a pathological hypervigilance frequently observed after traumatic experiences and a key factor for the emergence of multiple mental health disorders [@fonagy_mentalizing_2017].

@@

A well-established paradigm in the study of trust is the repeated trust game [@joyce_trust_1995]. To encourage the emergence and maintenance of trust in this setting, some studies focused on modifying the game mechanism [@charness_investment_2008; @fiedler_effect_2017]. Others chose to intervene directly on the participants. For example, @drazkowski_gratitude_2017 found that trust was increased when participants were asked to think about and write down five things that they were grateful for. @burnham_friend-or-foe_2000 found that trust was also increased when participants were primed with the concepts of friend and foe.

Whilst these interventions show that it is possible to improve cooperative outcomes at the start of the game, they did not address how to repair a breakdown of trust that might occur due to intentional or accidental non-cooperative actions by the players. Indeed, cooperative play in the repeated trust game can easily break down when there is a transgressive behavior, such as a nil or very low investment by the investor after cooperation has been established, or a return of the trustee below the investment sent. Such ruptures of cooperation appear frequently when the trustee suffers from mental health disorders affecting the social domain such as Borderline Personality Disorder [@lieb_borderline_2004]. In theses situations, BPD trustees fail to engage in trust repairing behavior such as coaxing the investor through sending high return to signal trustworthiness, and this failure may be linked to a failure to perceive their low returns in the game as a violation of social norms [@king-casas_rupture_2008].

In devising potential interventions to repair trust, we can derive inspiration from the cognitive interventions championed by successful psychological therapies that aim to improve aspects of interpersonal dysfunction in BPD patients. Whilst there is no proven pharmacological therapy for BPD, some forms of psychotherapy such as Mentalisation Based Therapy [@allen_handbook_2006-1] and Dialectical Behavior Therapy [@linehan_cognitive-behavioral_1993], have been clinically validated as efficacious approaches to improve various dysfunctional behaviors in BPD patients, including those related to social interaction [@gunderson_borderline_2018]. However, these therapies suffer from a high variability in treatment responses. Determining which interventions are effective for particular patients has been very challenging [@rudge_mechanisms_2020; @arch_longitudinal_2012]. One promising approach is the study of how specific psychotherapeutic treatment components affect quantitative markers of behavior such as those inferred through computational modelling techniques [@huys_computational_2016; @reiter_neuro-cognitive_2021]. As such, using specific cognitive probes inspired by therapeutic interventions complemented by computational modelling of behavior in tasks, may allow us to uncover cognitive mechanisms targeted by common forms of psychotherapy. This may in turn form the basis for targeting of existing psychotherapies to different individuals.

>>>>

```{r load-packages05, include = FALSE}
library(papaja)
library(kableExtra)
require(knitr)
#require(citr)
require(bookdown)

@@

Determining the other's goals, intentions, and decision-making process is fundamental to successful social interaction. This inference is however fraught with uncertainty, as we cannot directly observe these features, and may need extensive prior knowledge of the person. Absent a history of interaction, one may decide to trust the other's goals and intentions are aligned with one's own. The challenge of trust is that it is by construction a risky endeavour. If we deem a person trustworthy, we might take the risk of investing in the relationship, hoping for a collaborative outcome. If this trust is misplaced, it can come at a high cost. Not trusting others is also risky since opportunities for cooperation may be foregone.

@@

Evidence from the literature emphasises the importance of social trust in determining why some people fare better than others physically and mentally [@giordano_trust_2016 ; @meng_multilevel_2014]. It affects the health status of individuals by reinforcing social support networks, maintaining community norms and facilitating collective action. Research into the determinants of psychopathology has linked trust-based constructs to the emergence of mental health disorders. @fonagy_role_2014 identified epistemic trust, or the belief in the authenticity and personal relevance of interpersonally transmitted knowledge, as an important function of early attachment relationships. It allows the individual receiving social information to let go of their natural self-protective vigilance, which can become a pathological hypervigilance frequently observed after traumatic experiences and a key factor for the emergence of multiple mental health disorders [@fonagy_mentalizing_2017].

@@

A well-established paradigm in the study of trust is the repeated trust game [@joyce_trust_1995]. In this game, the "investor" decides how much of an endowment to send to the other player (the "trustee"). The amount that is sent is tripled and the trustee decides, in return, how much of the tripled amount to send back to the investor. To encourage the emergence and maintenance of trust in this setting, some studies focused on modifying the game mechanism, for example by introducing a third-party who monitors the actions of the other players [@charness_investment_2008; @fiedler_effect_2017]. Others chose to intervene directly on the participants. For example, @drazkowski_gratitude_2017 found that trust was increased when participants were asked to think about and write down five things that they were grateful for. @burnham_friend-or-foe_2000 found that trust was also increased when participants were primed with the concepts of friend and foe.

Whilst these interventions show that it is possible to improve cooperative outcomes at the start of the game, they do not address how to repair a breakdown of trust that might occur due to intentional or accidental non-cooperative actions by the players. Cooperative play in the repeated trust game can easily break down when there is a transgressive behavior, such as a nil or very low investment by the investor or a return of the trustee below the investment sent [@bendor_when_1991]. Such ruptures of cooperation appear frequently when the trustee suffers from social disorders such as Borderline Personality Disorder [BPD; @lieb_borderline_2004]. In theses situations, BPD trustees fail to engage in trust repairing behaviours such as coaxing the investor by signalling trustworthiness via sending high returns. This failure may be linked to the misperception that their low returns in the game are not violating social norms [@king-casas_rupture_2008].

In devising potential interventions to repair trust, we can derive inspiration from the cognitive interventions championed by successful psychological therapies that aim to improve aspects of social dysfunction in BPD patients. Whilst there is no proven pharmacological therapy for BPD, psychotherapies such as Mentalisation Based Therapy [@allen_handbook_2006-1] and Dialectical Behavior Therapy [@linehan_cognitive-behavioral_1993] are proven to improve various dysfunctional behaviors in BPD patients, including those related to social interaction [@gunderson_borderline_2018]. However, response to these treatments is highly variable, and determining which interventions are effective for particular patients is challenging [@rudge_mechanisms_2020; @arch_longitudinal_2012]. One promising approach is the study of how specific psychotherapeutic treatment components affect quantitative markers of behavior such as those inferred through computational modelling techniques [@huys_computational_2016; @reiter_neuro-cognitive_2021]. As such, using specific cognitive probes inspired by therapeutic interventions complemented by computational modelling of behavior in tasks, may allow us to uncover cognitive mechanisms targeted by common forms of psychotherapy. This may in turn form the basis for targeting of existing psychotherapies to different individuals.
